A Hexagon Tile Map Algorithm For Spatial Mapping To Create Alternative Australian Maps

### 2020-09-28 algorithm.tex
clean up for journal submission

---
title: A Hexagon Tile Map Algorithm For Spatial Mapping To Create Alternative Australian Maps
author:
  # see ?rjournal_article for more information
  - name: Stephanie Kobakian
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    email:  stephanie.kobakian@monash.edu 
  - name: Dianne Cook
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    email:  dicook@monash.edu 
abstract: >
  This algorithm creates a tessellated hexagon display to represent a set of the spatial polygons. It allocates these hexagon in a manner that preserves the spatial relationship of the geographic units. The algorithm creates a display that showcases spatial distributions, by emphasising the small geographical regions that are often difficult to locate on geographic maps. Spatial distributions have been presented on alternative representations of geography for many years. In modern times, interactivity and animation have begun to play a larger role, as alternative representations have been popularised by online news sites, and atlas websites with a focus on public consumption. Applications are increasingly widespread, especially in the areas of disease mapping, and election results.
preamble: |
  % Any extra LaTeX you need in the preamble
bibliography: sugarbag.bib
output: rticles::rjournal_article
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, comment = FALSE}
library(knitcitations)
library(RefManageR)
library(sf)
library(sugarbag)
library(kableExtra)

knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options("citation_format" = "pandoc")
BibOptions(check.entries = FALSE, style = "markdown")
```

## Introduction

<!-- Motivation to map -->
Many cancer atlases present geospatial cancer data on a choropleth map display.
The Australian Cancer Atlas [@TACA] is a recent addition to the many cancer atlas maps worldwide. The ground-breaking atlas for Australia presents a central map that shows the landmass overlaid with administrative boundaries. This choropleth display can highlight the geographic patterns in geospatially related cancer statistics [@SAMGIS].

<!-- Issues with choropleth map displays... -->
In many countries, residents are increasingly gathering to live in urban areas around major cities [@ACTUC].
The creation of approximately equal population areas results in dramatically different square meterage of the geographic units. The geographic units are usually administrative, such as states or electorates. The units are filled with colour to represent the value of the statistic for each unit [@EI].
This can lead to unequal emphasis on the statistical value of each of the geographic units. This allows choropleth map displays to misrepresent the spatial distributions of human related statistics due to area-size bias [@BCM].
Figure \ref{fig:melanoma-geo} is a choropleth map that uses colour to display the estimated Standardised Incidence Ratios (SIRs) of melanoma cancer for males for each of the Statistical Areas at Level 2 (SA2) used by the Australian Bureau of Statistics [-@abs2011]. The Australian choropleth map display draws attention to the expanse of light blue areas across the rural communities in all states. The SA2s around Brisbane stand out as more orange and red. To create this choropleth map the SA2 polygons for 2011 from the Australian Bureau of Statistics.

<!-- solutions -->
The solutions to this visualisation problem begin with the geography. Alternative maps shift the focus from land area and shape, to the value of the statistics in a group of areas [@ACCAC]. Different styles of cartograms apply transformations to the geographic areas, to highlight the values of the value of the statistic of interest. These displays result in a distortion of the map space to represent differences in the statistic across the areas [@ACCAC] as the statistic of interest is used to determine the cartogram layout.

@@

Alternative mapping methods allow increased understanding of the spatial distribution of a variable across the population, by fairly representing the population in each administrative area [@TAAM]. This acknowledges that the number of residents can be different but recognises that each area, or person within it is equally important.

This paper contains a discussion of Existing Mapping Practices in Section 2. This is followed by details of the Algorithm in Section 3. Section 4 describes the implementation of the algorithm in the `sugarbag` package. Section 5 outlines the use and benefits of animation.

## Existing Mapping Practices

There are several established alternative visualisation methods. 
Tile maps, Rectangular cartograms [@ORC] and Dorling cartograms [@ACTUC] all use one simple shape to represent each geographic unit. They all minimise the emphasis on the size or shape of the geographic areas. These alternative map displays focus on the relationship between neighbours, attempting to preserve connections, and disregard the unique shapes of the administrative boundaries. Figure \ref{fig:qld_grid} shows a collection of alternative map displays, this includes a) a contiguous cartogram, b) a non-contiguous cartogram, c) a Dorling cartogram and d) a hexagon tile map. 
Each Statistical Area at Level 2 (SA2) [@abs2011] was designed to represent a community. These administrative areas are often used to aggregate census data and used to understand demographics of the Australian population.

<!-- choropleth -->
The familiar choropleth map display is often used to allow the user to orient themselves. However, the emphasis on the land mass diminishes the features of the distribution in densely populated communities due to the small size on the display [@ACTUC].
The unique shapes of boundaries can be helpful for orienting users but may not contribute to their understanding of the spatial disease distribution as many of the communities are not visible in a choropleth display [@TVSSS]. This presents an opportunity for explicit map transformations to improve communication of distributions [@CBATCC]. In figure \ref{fig:qld_grid}a) the choropleth map is presented with the alternative maps to allow for comparisons to be made.

@@

<!-- contiguous -->
When communicating information that is relevant to the population, each member of the population can be given equal representation by transforming the map [@TVSSS].
The impact of a disease over connected communities can be seen by allowing the boundaries to maintain connection in the transformed display. The contiguous cartogram displayed in figure \ref{fig:qld_grid}b) draws attention to smaller geographic units when they are rescaled according to the population [@DMAHP]. These new shapes can now be coloured to represent a second variable. This display can create twisted and unfamiliar shapes from the geographic units as the algorithms must satisfy the topology conditions, especially when there are communities located geographically far from their neighbours [@TVSSS]. This shown in the northern SA2s of Queensland with an average (yellow) estimated SIR of Melanoma for males.

<!-- non-contiguous -->
The non-contiguous cartogram in figure \ref{fig:qld_grid}c) also uses the population to rescale the geographic units. Unlike the contiguous cartogram, the SA2 areas maintain their geographic shape, but they may not retain the connection with their neighbours. The population of the SA2 areas is used to scale the geographic units in the non-contiguous cartogram [@NAC]. The amount of white space can be meaningful in non-contiguous cartograms [@ECGC], in this example the disparity between large rural areas and small urban areas means that the city of Brisbane, shown in the inset map looks reasonable, but these units are not visible in the context of the Queensland state map. Depending on the difference between the population and geographic land size, the amount of white space can also prevent meaningful understanding of the distribution [@TVSSS].

<!-- Dorling -->
The Dorling cartogram presents each geographic unit as a circle, the size of the circle is scaled according to the population value of each area [@ACTUC]. Figure \ref{fig:qld_grid}d) shows the Queensland SA2 areas as an individual circle located as close as possible to the geographic centroid location. This map draws attention to the collection of coastal cities in Queensland that were not apparent in figure \ref{fig:qld_grid} a), b) or c).

<!-- When the population is the statistic of interest, the result is a population cartogram. They fail to preserve a recognisable display due to the difference in size of metropolitan and rural areas [@ACTUC], [@GOINO]. -->
<!-- Contiguous cartograms alter the shape of areas, while preserving boundary relationships of neighbours. Non-contiguous cartograms maintain the geographic shape of each geographic area, but will lose the connection to neighbours as the polygon for each geographic area shrinks or grows. -->

## Algorithm

@@

The algorithm is named for the *Trigona carbonaria* bee species. Native to northern and eastern Australia this stingless species builds flat layers of hexagonal brood cells, spiralling out from a central point [@PH]. This hive design inspired the use of multiple focal points in the algorithm, individual spirals are constructed outward from various points on the geographic map base.

@@

#### User choices

<!-- The`create_hexmap` function will -->
<!-- Derive the centroids of the geographic units to calculate a grid of possible hexagon locations, and allocate each geographic unit to a single hexagon. -->
Only two inputs are necessary to begin using the algorithm, the shape file, and the id variable. The id column should uniquely identify each geographic unit in the shape file.

@@

## Implementation

<!-- Motivate the value of the the package and algorithm -->
Hexagon tile maps can be useful to understand a distribution across a collection of geographic areas. However, these maps are not easily drawn, especially as the number of areas increases. This algorithm was created to automate this process, and reduce the manual load involved in creating and implementing alternative displays. This allows map makers and data communicators to allocate their time to choosing the most effective display, rather than manually creating them.

The `sugarbag` package has been written for R software, it contains a set of functions that help R users to create a hexagon tile map.
The algorithm presented in the `sugarbag` package operates on a set of simple feature geometry objects [@sf], this package allows R users to create `sf` objects by importing polygons stored in various formats.
Users should provide a set of polygons that define geographic units by their administrative boundaries. The functions arrange the geographic units in order of proximity to a set of locations provided, such as the centre of capital cities. 
The centroid location of each geographic unit is used to measure the proximity.
It emphasises the capital cities as population hubs, rather than emphasizing the size of large, rural geographic units.

A user can tweak the parameters of the hexagon map using the additional arguments to the `create_hexmap` function, these options may affect the speed of the creation.

#### Installation

@@

```{r eval=FALSE}
install.packages("sugarbag")
```

@@

```{r eval=FALSE}
devtools::install_github("srkobakian","sugarbag")
```

@@

```{r}
library(sugarbag)
```

#### Creating a hexagon tile map

The following code creates the hexagon tile map for all the Statistical Areas at Level 2 in Queensland. These steps can be executed by the main function, `create_hexmap`, or can be implemented separately for more flexibility.

@@

Users may choose to only use the `allocate` function when they wish to use a set of centroids, rather than [@sf] polygons.

```{r full_hexmapcode, echo = TRUE, warning = FALSE, message = FALSE, eval=FALSE}
## Load data
qld_sa2 <- absmapsdata::sa22011 %>% 
  filter(state_name_2011 == "Queensland")

@@

## Make a plot
library(ggplot2)
qld_hexmap <- ggplot() +
  geom_polygon(aes(
      x = long,
      y = lat,
      group = interaction(sa2_name_2011, polygon)
    ),
    fill = "white",
    colour = "lightgrey",
    data = fort_qld
  ) +
  geom_polygon(aes(
    x = long,
    y = lat,
    group = hex_id,
    fill = rownumber
  ),
  data = fort_hex) +
  scale_fill_distiller("", palette = "PRGn") +
  coord_equal() + theme_void()
qld_hexmap
```

@@

#### Polygon set

The polygon set of Statistical Areas at Level 2 (SA2) [@abs2011] of Queensland in 2011 is provided with the `sugarbag` package as `tas_sa2`. 
A single column of the data set is used to identify the unique areas. 
In this case, the unique SA2 names for each SA2 have been used.

@@

```{r capital_cities}
data(capital_cities)
```

The buffer distance, hexagon size, hexagon amount to filter and width of angle are parameters that will be determined within `create_hexmap` if they are not provided. They are created as they are needed throughout the following example.

#### Step 1: Derive the set of centroid points

A set of centroids may be used directly. The set of polygons should be provided as an `sf` object, this is a data frame containing a `geometry` column. The `read_shape` function can assist in creating this object for use in `R`.

The centroids can be derived from the set of polygons using the `create_centroids` function:

```{r cents}
centroids <- create_centroids(shp_sf = tas_sa2, sf_id = "SA2_NAME16")
```

@@

#### Step 2: Create the hexagon grid points

@@

The grid of possible hexagon locations is made using the `create_grid` function.
It uses the centroids, the hexagon size and the buffer distance.

```{r grid}
grid <- create_grid(centroids = centroids, hex_size = 0.2, buffer_dist = 1.2)
```

##### Creating a tessellated grid

A set of longitude columns, and latitude rows are created to define the locations of the hexagons. The distance between each row and column is the size specified by `hex_size`.
Equally spaced columns are created from the minimum longitude minus the buffer distance, up to the maximum longitude plus the buffer distance. Similarly, the rows are created from the latitude values and the buffer distance.
A unique hexagon location is created from all intersections of the longitude columns and latitude rows.
Figure \ref{fig:grid2} shows the original grid on the left, to allow for tessellating hexagons, every second latitude row on the grid is shifted right, by half of the hexagon size. The grid for tessellation is shown on the right.

<!-- ```{r grid2, fig.cap = "Grid points to create a tile map.", out.width = "100%", echo = FALSE} -->
<!-- knitr::include_graphics(path = "figures/2grid.pdf") -->
<!-- ``` -->

@@

##### Rolling windows

Not all of the grid points will be used, especially if islands result in a large grid space.
To filter the grid for appropriate hexagon locations for allocation, the `create_buffer` function is used by `create_grid`.
It finds the grid points needed to best capture the set of centroids on a hexagon tile map.

@@

##### Filtering the grid

@@

<!-- ```{r filter-grid, fig.cap = "All possible hexagon locations from the initial grid are shown with blue outlines. The blue dots show the grid points left to choose from after the buffer step. The rolling windows show the collections of rows and columns used to filter the hexagon locations.", echo = FALSE, out.width = "100%"} -->
<!-- knitr::include_graphics(path = "figures/3grid.pdf", dpi = 300) -->
<!-- ``` -->

@@

##### Centroid to focal point distance

@@

#### Step 3: Allocate each centroid to a hexagon grid point

@@

The possible hexagon locations to consider for a centroid are determined by the `hex_filter`. This is the maximum amount of hexagons between the centroid and the furthest considered hexagon. It is used to subset possible grid points to only those surrounding the polygon centroid within an appropriate range. A smaller distance will increase speed, but can decrease accuracy when width of the angle increases.

```{r centroids, eval = FALSE}
hexmap_allocation <- allocate(
  centroids = centroids %>% select(SA2_NAME16, longitude, latitude),
  sf_id = "SA2_NAME16",
  hex_grid = grid,
  hex_size = 0.2, ## same size used in create_grid
  hex_filter = 10,
  width = 35,
  focal_points = capital_cities,
  verbose = TRUE)
```

@@

##### Filter the grid for unassigned hexagon points

@@

##### Filter the grid points for those closest to the centroid

@@

<!-- ```{r buffers, fig.cap = "Filter for grid points within a square, then circular, distance for those closest to the centroid.", echo = FALSE, out.width = "100%"} -->
<!-- knitr::include_graphics(path = "figures/4grid.pdf") -->
<!-- ``` -->

@@

The `width` parameter is used to take a slice of the remaining points. The width is the amount of degrees used on either side of the angle from the focal point to centroid location. This uses the angle from the closest capital city, to the current centroid as seen in Figure \ref{fig:angles} . This allows the spatial relationship to be preserved, even when it is allocated to a hexagon that is further from the focal point then the original centroid location.

<!-- ```{r angles, fig.cap = "Filter for grid points within the angle from the focal point to the centroid.", echo = FALSE, out.width = "100%"} -->
<!-- knitr::include_graphics(path = "figures/5allocate.png") -->
<!-- ``` -->

@@

If no available hexagon grid point is found within the original filter distance and angle, the distance is expanded, only when a maximum distance is reached will the angle expand to accommodate more possible grid points.  
By default the angle filter to hexagon grid points that fall within the bounds of the angle from the focal point to the geographic centroid, plus and minus 30 degrees. This will increase if no points can be found within the `hex_filter` distance. The default angle of 30 was chosen to allow the algorithm to choose hexagons that best maintained the spatial relationship between the focal point and geographic centroid.

<!-- ```{r buffs, fig.cap = "A complete hexagon tile map of Queensland.", out.width = "100%", echo = FALSE} -->
<!-- knitr::include_graphics(path = "figures/6allocate.png") -->
<!-- ``` -->

@@

A complete hexagon tile map of Queensland is created by applying the algorithm steps to each centroid.
The hexagon tile map visualisation is used below to visualise the Australian Cancer Atlas data [@TACA]. Two views of the same data are produced by filling according to the Lung Cancer Standardised Incidence Rates (SIRs) downloaded from the Australian Cancer Atlas site. This small example in Figure \ref{fig:sir} shows the group of blue areas in the Brisbane CBD more prominently in the hexagon tile map (b).
The small red areas visible in the choropleth map (a) along the north coast are much larger in the hexagon tile maps. The hexagon tile map shows less yellow, this no longer overwhelms the map space with the information regarding the rural areas.

<!-- ```{r sir, out.width = "100%", echo = FALSE, fig.cap = "The Australian Cancer Atlas data has determined the colour of each Statistical Area of Australian at Level 2. A choropleth map (a) of Standardised Incidence Rates (SIRs) is paired with a hexagon tile map (b) to contrast the colours that are made obvious when every SA2 is equally represented."} -->
<!-- knitr::include_graphics(path = "figures/7SIR.png") -->
<!-- ``` -->

@@

#### Neighbour relationships

@@

An additional step can be included to allow the neighbours that have already been allocated to influence the placement of the current centroid. 
This requires specifying the `sf` object as the argument for the `use_neighbours` parameter. This calculates neighbours using intersections of their polygons. This occurs for all areas before any allocations begin.

@@

## Hexagon tile map of Australia

@@

## Animation

The `gganimate` [@gganimate] package can be used to make an animation.
It requires connecting the polygons for each area in two displays, which can be done using the `sf_id` variable, such as the SA2 name.
The animation^[This animation can be viewed at: https://sugarbagjss.netlify.com/] connecting these two displays will highlight the rapid growth of the inner-city areas that emphasises the density of the communities. The hexagons that move the furthest will move rapidly in the animation. The rapid decreases of the large rural areas also show how much greater the landmass of Statistical Areas can be.

## Conclusion  {#conclusion-03}

@@

Alternative maps highlight the value of the statistics across the geographic units. Alternative mapping methods allow increased understanding of the spatial distribution of a variable across the population, by fairly representing each administrative area. This acknowledges that the amount of residents can be different but recognises that each population area is equally important. The solution to this visualisation problem has equally sized areas, with neighbourhood boundary connections. This map algorithm is implemented in the `sugarbag` [@sugarbag] package written for `R` [@R].
The `sugarbag` package creates tessellated hexagon tile maps. The Australian application preserves the spatial relationships, emphasising capital cities. The hexagon tile map is a visualisation solution that highlights spatial distributions.

@@

<!-- future -->
Future work will include refining the algorithm. It would be possible to take a logarithmic function rather than a direct angle to help choose a closer hexagon to the original centroid location, before increasing the width of the angle used to filter the hexagons.

@@

>>>>

% !TeX root = RJwrapper.tex
\title{A Hexagon Tile Map Algorithm For Spatial Mapping To Create Alternative
Australian Maps}
\author{by Stephanie Kobakian and Dianne Cook}

\maketitle

\abstract{%
This algorithm creates a tessellated hexagon display to represent a set
of the spatial polygons. It allocates these hexagon in a manner that
preserves the spatial relationship of the geographic units. The
algorithm creates a display that showcases spatial distributions, by
emphasising the small geographical regions that are often difficult to
locate on geographic maps. Spatial distributions have been presented on
alternative representations of geography for many years. In modern
times, interactivity and animation have begun to play a larger role, as
alternative representations have been popularised by online news sites,
and atlas websites with a focus on public consumption. Applications are
increasingly widespread, especially in the areas of disease mapping, and
election results.
}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Many cancer atlases present geospatial cancer data on a choropleth map
display. The Australian Cancer Atlas \citep{TACA} is a recent addition
to the many cancer atlas maps worldwide. The ground-breaking atlas for
Australia presents a central map that shows the landmass overlaid with
administrative boundaries. This choropleth display can highlight the
geographic patterns in geospatially related cancer statistics
\citep{SAMGIS}.

In many countries, residents are increasingly gathering to live in urban
areas around major cities \citep{ACTUC}. The creation of approximately
equal population areas results in dramatically different square meterage
of the geographic units. The geographic units are usually
administrative, such as states or electorates. The units are filled with
colour to represent the value of the statistic for each unit \citep{EI}.
This can lead to unequal emphasis on the statistical value of each of
the geographic units. This allows choropleth map displays to
misrepresent the spatial distributions of human related statistics due
to area-size bias \citep{BCM}. Figure \ref{fig:melanoma-geo} is a
choropleth map that uses colour to display the estimated Standardised
Incidence Ratios (SIRs) of melanoma cancer for males for each of the
Statistical Areas at Level 2 (SA2) used by the Australian Bureau of
Statistics \citeyearpar{abs2011}. The Australian choropleth map display
draws attention to the expanse of light blue areas across the rural
communities in all states. The SA2s around Brisbane stand out as more
orange and red. To create this choropleth map the SA2 polygons for 2011
from the Australian Bureau of Statistics.

The solutions to this visualisation problem begin with the geography.
Alternative maps shift the focus from land area and shape, to the value
of the statistics in a group of areas \citep{ACCAC}. Different styles of
cartograms apply transformations to the geographic areas, to highlight
the values of the value of the statistic of interest. These displays
result in a distortion of the map space to represent differences in the
statistic across the areas \citep{ACCAC} as the statistic of interest is
used to determine the cartogram layout.

@@

Alternative mapping methods allow increased understanding of the spatial
distribution of a variable across the population, by fairly representing
the population in each administrative area \citep{TAAM}. This
acknowledges that the number of residents can be different but
recognises that each area, or person within it is equally important.

This paper contains a discussion of Existing Mapping Practices in
Section 2. This is followed by details of the Algorithm in Section 3.
Section 4 describes the implementation of the algorithm in the
\texttt{sugarbag} package. Section 5 outlines the use and benefits of
animation.

\hypertarget{existing-mapping-practices}{%
\subsection{Existing Mapping
Practices}\label{existing-mapping-practices}}

There are several established alternative visualisation methods. Tile
maps, Rectangular cartograms \citep{ORC} and Dorling cartograms
\citep{ACTUC} all use one simple shape to represent each geographic
unit. They all minimise the emphasis on the size or shape of the
geographic areas. These alternative map displays focus on the
relationship between neighbours, attempting to preserve connections, and
disregard the unique shapes of the administrative boundaries. Figure
\ref{fig:qld_grid} shows a collection of alternative map displays, this
includes a) a contiguous cartogram, b) a non-contiguous cartogram, c) a
Dorling cartogram and d) a hexagon tile map. Each Statistical Area at
Level 2 (SA2) \citep{abs2011} was designed to represent a community.
These administrative areas are often used to aggregate census data and
used to understand demographics of the Australian population.

The familiar choropleth map display is often used to allow the user to
orient themselves. However, the emphasis on the land mass diminishes the
features of the distribution in densely populated communities due to the
small size on the display \citep{ACTUC}. The unique shapes of boundaries
can be helpful for orienting users but may not contribute to their
understanding of the spatial disease distribution as many of the
communities are not visible in a choropleth display \citep{TVSSS}. This
presents an opportunity for explicit map transformations to improve
communication of distributions \citep{CBATCC}. In figure
\ref{fig:qld_grid}a) the choropleth map is presented with the
alternative maps to allow for comparisons to be made.

@@

When communicating information that is relevant to the population, each
member of the population can be given equal representation by
transforming the map \citep{TVSSS}. The impact of a disease over
connected communities can be seen by allowing the boundaries to maintain
connection in the transformed display. The contiguous cartogram
displayed in figure \ref{fig:qld_grid}b) draws attention to smaller
geographic units when they are rescaled according to the population
\citep{DMAHP}. These new shapes can now be coloured to represent a
second variable. This display can create twisted and unfamiliar shapes
from the geographic units as the algorithms must satisfy the topology
conditions, especially when there are communities located geographically
far from their neighbours \citep{TVSSS}. This shown in the northern SA2s
of Queensland with an average (yellow) estimated SIR of Melanoma for
males.

The non-contiguous cartogram in figure \ref{fig:qld_grid}c) also uses
the population to rescale the geographic units. Unlike the contiguous
cartogram, the SA2 areas maintain their geographic shape, but they may
not retain the connection with their neighbours. The population of the
SA2 areas is used to scale the geographic units in the non-contiguous
cartogram \citep{NAC}. The amount of white space can be meaningful in
non-contiguous cartograms \citep{ECGC}, in this example the disparity
between large rural areas and small urban areas means that the city of
Brisbane, shown in the inset map looks reasonable, but these units are
not visible in the context of the Queensland state map. Depending on the
difference between the population and geographic land size, the amount
of white space can also prevent meaningful understanding of the
distribution \citep{TVSSS}.

The Dorling cartogram presents each geographic unit as a circle, the
size of the circle is scaled according to the population value of each
area \citep{ACTUC}. Figure \ref{fig:qld_grid}d) shows the Queensland SA2
areas as an individual circle located as close as possible to the
geographic centroid location. This map draws attention to the collection
of coastal cities in Queensland that were not apparent in figure
\ref{fig:qld_grid} a), b) or c).

\hypertarget{algorithm}{%
\subsection{Algorithm}\label{algorithm}}

@@

The algorithm is named for the \emph{Trigona carbonaria} bee species.
Native to northern and eastern Australia this stingless species builds
flat layers of hexagonal brood cells, spiralling out from a central
point \citep{PH}. This hive design inspired the use of multiple focal
points in the algorithm, individual spirals are constructed outward from
various points on the geographic map base.

@@

\hypertarget{user-choices}{%
\paragraph{User choices}\label{user-choices}}

Only two inputs are necessary to begin using the algorithm, the shape
file, and the id variable. The id column should uniquely identify each
geographic unit in the shape file.

@@

\hypertarget{implementation}{%
\subsection{Implementation}\label{implementation}}

Hexagon tile maps can be useful to understand a distribution across a
collection of geographic areas. However, these maps are not easily
drawn, especially as the number of areas increases. This algorithm was
created to automate this process, and reduce the manual load involved in
creating and implementing alternative displays. This allows map makers
and data communicators to allocate their time to choosing the most
effective display, rather than manually creating them.

The \texttt{sugarbag} package has been written for R software, it
contains a set of functions that help R users to create a hexagon tile
map. The algorithm presented in the \texttt{sugarbag} package operates
on a set of simple feature geometry objects \citep{sf}, this package
allows R users to create \texttt{sf} objects by importing polygons
stored in various formats. Users should provide a set of polygons that
define geographic units by their administrative boundaries. The
functions arrange the geographic units in order of proximity to a set of
locations provided, such as the centre of capital cities. The centroid
location of each geographic unit is used to measure the proximity. It
emphasises the capital cities as population hubs, rather than
emphasizing the size of large, rural geographic units.

A user can tweak the parameters of the hexagon map using the additional
arguments to the \texttt{create\_hexmap} function, these options may
affect the speed of the creation.

\hypertarget{installation}{%
\paragraph{Installation}\label{installation}}

@@

\begin{Schunk}
\begin{Sinput}
install.packages("sugarbag")
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
devtools::install_github("srkobakian","sugarbag")
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
library(sugarbag)
\end{Sinput}
\end{Schunk}

\hypertarget{creating-a-hexagon-tile-map}{%
\paragraph{Creating a hexagon tile
map}\label{creating-a-hexagon-tile-map}}

The following code creates the hexagon tile map for all the Statistical
Areas at Level 2 in Queensland. These steps can be executed by the main
function, \texttt{create\_hexmap}, or can be implemented separately for
more flexibility.

@@

Users may choose to only use the \texttt{allocate} function when they
wish to use a set of centroids, rather than \citep{sf} polygons.

\begin{Schunk}
\begin{Sinput}
## Load data
qld_sa2 <- absmapsdata::sa22011 %>% 
  filter(state_name_2011 == "Queensland")

@@

## Make a plot
library(ggplot2)
qld_hexmap <- ggplot() +
  geom_polygon(aes(
      x = long,
      y = lat,
      group = interaction(sa2_name_2011, polygon)
    ),
    fill = "white",
    colour = "lightgrey",
    data = fort_qld
  ) +
  geom_polygon(aes(
    x = long,
    y = lat,
    group = hex_id,
    fill = rownumber
  ),
  data = fort_hex) +
  scale_fill_distiller("", palette = "PRGn") +
  coord_equal() + theme_void()
qld_hexmap
\end{Sinput}
\end{Schunk}

@@

\hypertarget{polygon-set}{%
\paragraph{Polygon set}\label{polygon-set}}

The polygon set of Statistical Areas at Level 2 (SA2) \citep{abs2011} of
Queensland in 2011 is provided with the \texttt{sugarbag} package as
\texttt{tas\_sa2}. A single column of the data set is used to identify
the unique areas. In this case, the unique SA2 names for each SA2 have
been used.

@@

\begin{Schunk}
\begin{Sinput}
data(capital_cities)
\end{Sinput}
\end{Schunk}

The buffer distance, hexagon size, hexagon amount to filter and width of
angle are parameters that will be determined within
\texttt{create\_hexmap} if they are not provided. They are created as
they are needed throughout the following example.

\hypertarget{step-1-derive-the-set-of-centroid-points}{%
\paragraph{Step 1: Derive the set of centroid
points}\label{step-1-derive-the-set-of-centroid-points}}

A set of centroids may be used directly. The set of polygons should be
provided as an \texttt{sf} object, this is a data frame containing a
\texttt{geometry} column. The \texttt{read\_shape} function can assist
in creating this object for use in \texttt{R}.

The centroids can be derived from the set of polygons using the
\texttt{create\_centroids} function:

\begin{Schunk}
\begin{Sinput}
centroids <- create_centroids(shp_sf = tas_sa2, sf_id = "SA2_NAME16")
\end{Sinput}
\end{Schunk}

@@

\hypertarget{step-2-create-the-hexagon-grid-points}{%
\paragraph{Step 2: Create the hexagon grid
points}\label{step-2-create-the-hexagon-grid-points}}

@@

The grid of possible hexagon locations is made using the
\texttt{create\_grid} function. It uses the centroids, the hexagon size
and the buffer distance.

\begin{Schunk}
\begin{Sinput}
grid <- create_grid(centroids = centroids, hex_size = 0.2, buffer_dist = 1.2)
\end{Sinput}
\end{Schunk}

\hypertarget{creating-a-tessellated-grid}{%
\subparagraph{Creating a tessellated
grid}\label{creating-a-tessellated-grid}}

A set of longitude columns, and latitude rows are created to define the
locations of the hexagons. The distance between each row and column is
the size specified by \texttt{hex\_size}. Equally spaced columns are
created from the minimum longitude minus the buffer distance, up to the
maximum longitude plus the buffer distance. Similarly, the rows are
created from the latitude values and the buffer distance. A unique
hexagon location is created from all intersections of the longitude
columns and latitude rows. Figure \ref{fig:grid2} shows the original
grid on the left, to allow for tessellating hexagons, every second
latitude row on the grid is shifted right, by half of the hexagon size.
The grid for tessellation is shown on the right.

@@

\hypertarget{rolling-windows}{%
\subparagraph{Rolling windows}\label{rolling-windows}}

Not all of the grid points will be used, especially if islands result in
a large grid space. To filter the grid for appropriate hexagon locations
for allocation, the \texttt{create\_buffer} function is used by
\texttt{create\_grid}. It finds the grid points needed to best capture
the set of centroids on a hexagon tile map.

@@

\hypertarget{filtering-the-grid}{%
\subparagraph{Filtering the grid}\label{filtering-the-grid}}

@@

@@

\hypertarget{centroid-to-focal-point-distance}{%
\subparagraph{Centroid to focal point
distance}\label{centroid-to-focal-point-distance}}

@@

\hypertarget{step-3-allocate-each-centroid-to-a-hexagon-grid-point}{%
\paragraph{Step 3: Allocate each centroid to a hexagon grid
point}\label{step-3-allocate-each-centroid-to-a-hexagon-grid-point}}

@@

The possible hexagon locations to consider for a centroid are determined
by the \texttt{hex\_filter}. This is the maximum amount of hexagons
between the centroid and the furthest considered hexagon. It is used to
subset possible grid points to only those surrounding the polygon
centroid within an appropriate range. A smaller distance will increase
speed, but can decrease accuracy when width of the angle increases.

\begin{Schunk}
\begin{Sinput}
hexmap_allocation <- allocate(
  centroids = centroids %>% select(SA2_NAME16, longitude, latitude),
  sf_id = "SA2_NAME16",
  hex_grid = grid,
  hex_size = 0.2, ## same size used in create_grid
  hex_filter = 10,
  width = 35,
  focal_points = capital_cities,
  verbose = TRUE)
\end{Sinput}
\end{Schunk}

@@

\hypertarget{filter-the-grid-for-unassigned-hexagon-points}{%
\subparagraph{Filter the grid for unassigned hexagon
points}\label{filter-the-grid-for-unassigned-hexagon-points}}

@@

\hypertarget{filter-the-grid-points-for-those-closest-to-the-centroid}{%
\subparagraph{Filter the grid points for those closest to the
centroid}\label{filter-the-grid-points-for-those-closest-to-the-centroid}}

@@

@@

The \texttt{width} parameter is used to take a slice of the remaining
points. The width is the amount of degrees used on either side of the
angle from the focal point to centroid location. This uses the angle
from the closest capital city, to the current centroid as seen in Figure
\ref{fig:angles} . This allows the spatial relationship to be preserved,
even when it is allocated to a hexagon that is further from the focal
point then the original centroid location.

@@

If no available hexagon grid point is found within the original filter
distance and angle, the distance is expanded, only when a maximum
distance is reached will the angle expand to accommodate more possible
grid points.\\
By default the angle filter to hexagon grid points that fall within the
bounds of the angle from the focal point to the geographic centroid,
plus and minus 30 degrees. This will increase if no points can be found
within the \texttt{hex\_filter} distance. The default angle of 30 was
chosen to allow the algorithm to choose hexagons that best maintained
the spatial relationship between the focal point and geographic
centroid.

@@

A complete hexagon tile map of Queensland is created by applying the
algorithm steps to each centroid. The hexagon tile map visualisation is
used below to visualise the Australian Cancer Atlas data \citep{TACA}.
Two views of the same data are produced by filling according to the Lung
Cancer Standardised Incidence Rates (SIRs) downloaded from the
Australian Cancer Atlas site. This small example in Figure \ref{fig:sir}
shows the group of blue areas in the Brisbane CBD more prominently in
the hexagon tile map (b). The small red areas visible in the choropleth
map (a) along the north coast are much larger in the hexagon tile maps.
The hexagon tile map shows less yellow, this no longer overwhelms the
map space with the information regarding the rural areas.

@@

\hypertarget{neighbour-relationships}{%
\paragraph{Neighbour relationships}\label{neighbour-relationships}}

@@

An additional step can be included to allow the neighbours that have
already been allocated to influence the placement of the current
centroid. This requires specifying the \texttt{sf} object as the
argument for the \texttt{use\_neighbours} parameter. This calculates
neighbours using intersections of their polygons. This occurs for all
areas before any allocations begin.

@@

\hypertarget{hexagon-tile-map-of-australia}{%
\subsection{Hexagon tile map of
Australia}\label{hexagon-tile-map-of-australia}}

@@

\hypertarget{animation}{%
\subsection{Animation}\label{animation}}

The \texttt{gganimate} \citep{gganimate} package can be used to make an
animation. It requires connecting the polygons for each area in two
displays, which can be done using the \texttt{sf\_id} variable, such as
the SA2 name. The animation\footnote{This animation can be viewed at:
  \url{https://sugarbagjss.netlify.com/}} connecting these two displays
will highlight the rapid growth of the inner-city areas that emphasises
the density of the communities. The hexagons that move the furthest will
move rapidly in the animation. The rapid decreases of the large rural
areas also show how much greater the landmass of Statistical Areas can
be.

\hypertarget{conclusion-03}{%
\subsection{Conclusion}\label{conclusion-03}}

@@

Alternative maps highlight the value of the statistics across the
geographic units. Alternative mapping methods allow increased
understanding of the spatial distribution of a variable across the
population, by fairly representing each administrative area. This
acknowledges that the amount of residents can be different but
recognises that each population area is equally important. The solution
to this visualisation problem has equally sized areas, with
neighbourhood boundary connections. This map algorithm is implemented in
the \texttt{sugarbag} \citep{sugarbag} package written for \texttt{R}
\citep{R}. The \texttt{sugarbag} package creates tessellated hexagon
tile maps. The Australian application preserves the spatial
relationships, emphasising capital cities. The hexagon tile map is a
visualisation solution that highlights spatial distributions.

@@

Future work will include refining the algorithm. It would be possible to
take a logarithmic function rather than a direct angle to help choose a
closer hexagon to the original centroid location, before increasing the
width of the angle used to filter the hexagons.

@@

\bibliography{sugarbag.bib}
